### 2017-07-22 old/inst/doc/msde-exmodels.Rmd
multiple models tested

---
title: "Example SDE models provided by **`msde`**"
author: "Martin Lysy, JunYong Tong, Nigel Delaney"
date: "`r Sys.Date()`"
output:
  html_vignette:
    toc: true
bibliography: references.bib
csl: taylor-and-francis-harvard-x.csl
link-citations: true
vignette: >
  %\VignetteIndexEntry{Example models provided by msde}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

\newcommand{\bm}[1]{\boldsymbol{#1}}
\newcommand{\tth}{\bm{\theta}}
\newcommand{\pphi}{\bm{\phi}}
\newcommand{\eeta}{\bm{\eta}}
\newcommand{\Y}{\bm{Y}}
\newcommand{\y}{\bm{y}}
\newcommand{\B}{\bm{B}}
\renewcommand{\a}{\alpha}
\renewcommand{\b}{\beta}
\newcommand{\g}{\gamma}
\newcommand{\rv}[3][1]{#2_{#1},\ldots,#2_{#3}}
\newcommand{\dr}{\bm{\Lambda}}
\newcommand{\df}{\bm{\Sigma}}
\newcommand{\dt}{\Delta t}
\newcommand{\ud}{\mathrm{d}}
\newcommand{\var}{\mathrm{var}}
\newcommand{\cov}{\mathrm{cov}}
\newcommand{\cor}{\mathrm{cor}}
\newcommand{\N}{\mathcal{N}}
\newcommand{\Ym}{\Y_{\mathrm{miss}}}
\newcommand{\iid}{\stackrel{\mathrm{iid}}{\sim}}
\newcommand{\ind}{\stackrel{\mathrm{ind}}{\sim}}
\newcommand{\GGam}{\bm{\Gamma}}
\newcommand{\LLam}{\bm{\lambda}}
\newcommand{\PPsi}{\bm{\Psi}}

This vignette contains a complete description of the sample models found in `msde::example.models()`.

## Heston's stochastic volatility model

Let $S_t$ denote the value of a financial asset at time $t$.  Heston's stochastic volatility model [@heston93] is given by the pair of stochastic differential equations
$$
\begin{split}
\ud S_t & = \alpha S_t\ud t + V_t^{1/2}S_t\ud B_{1t} \\
\ud V_t & = -\gamma(V_t - \mu)\ud t + \sigma V_t^{1/2} \ud B_{2t},
\end{split}
$$
where $V_t$ is a latent stochastic volatility process, and $B_{1t}$ and $B_{2t}$ are Brownian motions with $\cor(B_{1t}, B_{2t}) = \rho$.  To improve the accuracy of the numerical discretization scheme used for inference, `msde::example.models` considers the transformed variables $X_t = \log(S_t)$ and $Z_t = 2 V_t^{1/2}$, for which Heston's SDE becomes
$$
\begin{split}
\ud X_t & = (\alpha - \tfrac 1 8 Z_t^2)\ud t + \tfrac 1 2 Z_t \ud B_{1t} \\
\ud Z_t & = (\beta/Z_t - \tfrac \gamma 2 Z_t)\ud t + \sigma \ud B_{2t},
\end{split}
$$
with $\cor(B_{1t}, B_{2t}) = \rho$.  Thus the diffusion function on the variance scale is
$$
\df_\tth(\Y_t) = \begin{bmatrix} \tfrac 1 4 Z_t^2 & \tfrac \sigma 2 Z_t \\ \tfrac \sigma 2 Z_t & \sigma^2 \end{bmatrix},
$$
where $\Y_t = (X_t, Z_t)$ and $\tth = (\alpha, \gamma, \beta, \sigma, \rho)$.  The data and parameter restrictions are $Z_t, \gamma, \sigma > 0$, $|\rho| < 1$, and $\beta > \tfrac 1 2 \sigma^2$, with the final restriction ensuring that $Z_t > 0$ with probability 1.  This model is contained in `msde::example.models(model = "hest")`.

## Bivariate Ornstein-Uhlenbeck process

This model for $\Y_t = (X_{1t}, X_{2t})$ is given by
$$
\ud \Y_t = (\GGam \Y_t + \LLam)\ud t + \PPsi \ud \bm{B}_t,
$$
where $\GGam$ is a $2\times 2$ matrix, $\LLam$ is a $2 \times 1$ vector, and $\PPsi$ is a $2\times 2$ upper Choleski factor.  The model parameters are thus $\tth = (\Gamma_{11}, \Gamma_{21}, \Gamma_{12}, \Gamma_{22}, \lambda_{1}, \lambda_2, \Psi_{11}, \Psi_{21}, \Psi_{22})$, and the model restrictions are $\Psi_{11}, \Psi_{22} > 0$.  This model is contained in `msde::example.models(model = "biou")`.

## Lotka-Volterra predator-prey model

Let $H_t$ and $L_t$ denote the number of Hare and Lynx at time $t$ coexisting in a given habitat.  The Lotka-Volterra SDE describing the interactions between these two animal populations is given by [@golightly-wilkinson10]:
$$
\begin{bmatrix} \mathrm{d} H_t \\ \mathrm{d} L_t \end{bmatrix} = \begin{bmatrix} \a H_t - \b H_tL_t \\ \b H_tL_t - \g L_t \end{bmatrix}\, \mathrm{d} t + \begin{bmatrix} \a H_t + \b H_tL_t & -\b H_tL_t \\ -\b H_tL_t & \b H_tL_t + \g L_t\end{bmatrix}^{1/2} \begin{bmatrix} \mathrm{d} B_{1t} \\ \mathrm{d} B_{2t} \end{bmatrix}.
$$
The data and parameters are all restricted to be positive.  This model is contained in `msde::example.models(model = "lotvol")`.

## Prokaryotic autoregulatory gene network model

Let $\Y_t = (R_t, P_t, Q_t, D_t)$ denote the number of molecules at time $t$ of four different compounds in an autoregulatory gene network: RNA ($R$); a functional protein ($P$); protein dimmers ($Q$); and DNA ($D$). Then @golightly.wilkinson05 define an SDE describing the dynamics of $\Y_t$ with drift and (variance-scale) diffusion functions
$$
\begin{split}
\dr_\tth(\Y_t) & =
\begin{bmatrix}
\gamma_3 D_t - \gamma_7 R_t \\
2 \gamma_6 Q_t - \gamma_8P_t + \gamma_4 R_t -\gamma_5 P_t(P_t-1) \\
\gamma_2(10-D_t) - \gamma_1 D_t Q_t - \gamma_6 Q_t + \tfrac 1 2 \gamma_6 P_t(P_t-1) \\
\gamma_2(10-D_t) - \gamma_1 D_t Q_t
\end{bmatrix} \\
\df_\tth(\Y_t) & =
\begin{bmatrix}
\gamma_3 D_t + \gamma_7 R_t & 0 & 0 & 0 \\
0 &  \gamma_8P_t + 4\gamma_6 Q_t + \gamma_4 R_t + 2 \gamma_5 P_t(P_t-1) & -2 \gamma_6 Q_t - \gamma_5 P_t(P_t-1) & 0 \\
0 & -2 \gamma_6 Q_t - \gamma_5 P_t(P_t-1) & A + \gamma_6 Q_t + \tfrac 1 2 \gamma_5 P_t(P_t-1) & A_t \\
0 & 0 & A_t & A_t
\end{bmatrix},
\end{split}
$$
where $A_t = \gamma_1D_tQ_t + \gamma_2(10-D_t)$ and $\tth = (\rv \theta 8)$, $\theta_i = \log(\gamma_i)$, are various reaction rates.  The data and parameter restrictions for this model are $\tth \in \mathbb R^8$, $\Y_t > 1$, and $D_t < 10$.  This model is contained in `msde::example.models(model = "pgnet")`.

## References