### 2021-12-09 tests/dontrun/Rcpp_templates.Rmd
implemented OOP via RC, passes Rdevel CMD check --use-valgrind

---
title: "C++ class templates with Rcpp and R6"
author: "Martin Lysy"
output: rmarkdown::html_document
---

```{r setup, include = FALSE}
cran_link <- function(pkg) {
  paste0("[**", pkg, "**](https://CRAN.R-project.org/package=", pkg, ")")
}
```

# Motivation

Consider the following paradigm for writing methods for parameter inference that can be applied to any user-defined member of a family of models.  This is achieved using a templated class `Inference<Model>` of which the template parameter `Model` defines model-specific versions of methods which `Inference` might use (e.g., `Model.loglik()`).  Within the proposed paradigm we'll make the following assumption:

> The only public members of the `Inference<Model>` class (that get exported to R) all have the same signature, including the class constructor, and all have arguments and return values than can be wrapped by `r cran_link("Rcpp")`.

So for example:
```cpp
/// User-defined model.
class IIDNormal {
 public:
  double loglik(const NumericVector& theta, const NumericVector& x) {
	double ll = 0.0;
	for(int ii=0; ii<x.length(); ii++) {
	  ll += R:dnorm(x(ii), theta(0), theta(1), true);
	}
	return ll;
  }
};

/// Generic methods for inference.
template <class Model>
class Inference {
 private:
  Model model_;
 public:
  // Calculate mle based on x and model.loglik(theta, x)
  void mle(NumericVector& theta_hat, const NumericVector);
};

// C++ API
iidnorm_fit = Inference<IIDNormal>; // create inference object
iidnorm_fit.mle(theta_hat, x); // calculate MLE
```

The objective is to create an R wrapper to `Inference<IIDNormal>` and its public methods.  It is proposed to do this using **Rcpp** and `r cran_link("R6")` classes as follows.

# **Rcpp** Wrapper

We could use **Rcpp** [modules](https://dirk.eddelbuettel.com/code/rcpp/Rcpp-modules.pdf) for this.  However, there are a few caveats to this approach:

- The R wrapper is an RC class, not R6, but the latter is much more lightweight, flexible, and easy to use.

- The R wrapper provides methods which accept arguments positionally and with zero error checking.  This is in contrast to how **Rcpp** would export the following function:

	```cpp
	//[[Rcpp::export]]
	double foo(double x, NumericVector y)
	```

    That is, the R call `foo(y = 1:5, x = 3)` would work as expected and `foo(1:5, x)` would give an error, whereas `obj$foo()` as a class method exported by **Rcpp** modules would in both cases crash the R session.

Bearing this in mind, the approach advocated here is to create something like `Inference_IIDNormal_Exports.cpp` which wraps the class functionality in a rudimentary but transparent way.  Since we don't want to do this manually for each `Model`, we can use the [Mustache](https://mustache.github.io/) templating logic implemented in the R package `r cran_link("whisker")`.  So for example the template `cpp` file would look like this:

```
/// @file Inference_{{Model}}_Exports.cpp
///
/// Rcpp wrapper for the methods of Inference<{{Model}}>

#include {{Model_file}} // {{Model}} class definition
#include Inference.h // Inference class definition

//[[Rcpp::export]]

```

# R6 Wrapper

On the R side, we have an R6 class called `Inference_IIDNormal`.